Report on Late Winter2013 Production: Image Differencing\vspace{-2ex}

### 2013-04-01 W13report.tex
Final comments and clarifications on MSE and threshold parts

We can model source free image (i.e.\ with a background described by a
Gaussian process) using the statistics of Gaussian random fields
(Kaiser 2004). For an image with Gaussian noise convolved with a
Gaussian PSF with width $\sigma_g$ the density of peaks above a given
detection threshold, $\nu$ is given by 
\begin{equation}
n(>\nu) = \frac{1}{2^{5/2}\pi^{3/2}\sigma_g^{2}} \nu e^{-\nu^2 /2} 
\end{equation}
 To estimate the number of random
positive detections for a given image we must normalize this density
by the total number of realizations of the PSF within the image, i.e.,
\begin{equation}
N_{total}(>\nu) = n(\nu)*nrow*ncol/PSF_{FWHM}^2
\end{equation}
where nrow and ncol are the number of rows and columns in an images
and $PSF_{FWHM}$ is the FWHM of the PSF.

Figure~\ref{fig:peaks} shows good agreement in a comparison of this
model to a Monte-Carlo simulation for thresholds $4<\nu<6$ for a
4000x4072 pixel image with 0.88 arcsec seeing. The variance on these
measures is shown in Figure XXX
\begin{figure}
\includegraphics[width=0.5\textwidth,
height=0.25\textwidth]{figures/peaks.eps} 
\caption{A comparison of the predicted number of random peaks (solid
  line) detected in an image, as a function of significance, with
  those derived from a Monte-Carlo realization of a Gaussian noise
  image (solid points).}
\label{fig:peaks}
\end{figure}

@@

\subsection{Detection Threshold}
Obviously the number of false positives is strongly dependent on the detection threshold.  In order to
check the scaling of our false positive rates with detection threshold, we ran the same config (see Table~\ref{tab-bestconfig10} 
for the config parameters)
at several detection threshold values.  As discussed in 
Section~\ref{sec-analyticfp} one can predict the number of false positives expected in a sensor given
seeing and threshold values.  Values of the number of false positives (both +ve and -ve)
expected in LSST size detectors are given in Table~\ref{tab-fp}.

@@

In all cases the empirical results are below the predicted results indicating
that the number counts are smaller than the theoretical floor.  This is not entirely unexpected, as each
curve is strongly correlated, being a cumulative distribution, and represents only 9 samplings of the
background random Gaussian field.  We find that we can perform a 1--parameter correction to these curves that
yields both the same slope and zeropoints as the theoretical curves, with the exception of the deconvolution 
visit.  This correction involves scaling {\it down} the variance in the images by 2.5\%, 1.5\% and 1\% for 
v6866601, v6866602, and v6866603 respectively.  The data plotted with these corrections are shown
in the bottom row of Figure\ref{fig-fp_v_thresh}.  These results suggest our variance estimates 
are too {\it high}, and we are therefore detecting too {\it few} false detections.  This does not agree 
entirely with the analysis of the empirical variance, Section~\ref{sec-varmis} and Table~\ref{tab-variance1}, 
which suggest that our variance planes {\it underestimate} the true variance.  However, this makes clear that 
while we are at the level of the theoretical floor in false detection rate, it is imperative to track the 
variance in our data to at the 1\% level or better to optimize both the false detection rate, and the true detection
efficiency.

@@

Using this simplified data set, we can begin to look at how we may predict 
optimal configs based on measured statistics.  In particular, we investigate
the MSE, and show how it varies as a function of config in Figures~\ref{fig-MSE_FITTED} and \ref{fig-MSE_CONTROL}.
We find that the MSE of the fitted sample (Figure~\ref{fig-MSE_FITTED}) continually decreases as the complexity 
of the model increases.  However, for the control sample (Figure~\ref{fig-MSE_CONTROL}), a minimum is achieved in all cases.
This minimum value suggests the point where the model starts overfitting, and is itself a method
to suggest the optimal configuration for image differencing, along with false positive rate.
However, MSE of a control sample is a metric that may be applied to data where the truth is not known, and thus has
potential to be used in the nightly processing.
For more discussion on this see Section $8.11$ in
``The Second Best Book in the World'' (i.e. {Ivezi{\'c} et al., in press).

@@

\begin{table}
\centering
\begin{tabular}{clcccc}
\hline
\multicolumn{6}{|c|}{Per Sensor MSE Comparison} \\
\hline
Visit    & Prefilter & N$_{Same}$ & N$_{<3}$ & $\Delta FP$ & max($\Delta FP$)\\
\hline
v6866601 & True      & 0      & 2        & 22 & 4\\
         & False     & 4      & 6        & 24 & 10\\
v6866602 & True      & 3      & 5        & 7 & 2\\
         & False     & 6      & 7        & 8 & 6\\
v6866603 & True      & 6      & 7        & 3 & 1\\
         & False     & 6      & 8        & 3 & 1\\
\end{tabular}
\caption{
We examine the configs chosen based on the minimum MSE and compare them to the config of the
minimum number of false positives.  The sample here consists of 9 sensors per visit.  
N$_{Same}$ is the number of times the minimum MSE coincides with the minimum in false positives.  N$_{<3}$
is the number of times the minimum false positive config is within the best 3 of MSE.  $\Delta FP$
is the total change in false positive count using the minimum MSE config as compared to the sum of the minimum 
false positives per sensor.  Column max($\Delta FP$) is the maximum change in detected false positives on any sensor in that visit.
\label{tab-MSE_persensor}}
\end{table}

@@

\begin{figure}%
\captionsetup[subfloat]{labelformat=empty}
\centering
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_01_nocorr.eps}}
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_02_nocorr.eps}}
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_03_nocorr.eps}}\\
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_01_corr.eps}}%
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_02_corr.eps}}%
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_03_corr.eps}}%
\caption{Here we show the comparison of detected false positive count to the analytic prediction.
In all cases the blue circles are the prefilter case and the red squares are the postfilter case.  
The top row has had the counts corrected for the underestimate of the noise by the variance plane (Table~\ref{tab-variance1}).  
In the bottom row we undertaken a 1--parameter correction to align the predicted and measured values.  From left to right
this is requires a 5\%, 3\%, 2\% reduction of the image variance, respectively.  
The shaded area is the 1-$\sigma$ confidence envelope
determined from Monte Carlo simulations assuming the analytic form of the false detection rate and a Gaussian distribution.}
\label{fig-fp_v_thresh}
\end{figure}

>>>>

We can model source free image (i.e.\ with a background described by a
Gaussian process) using the statistics of Gaussian random fields
(Kaiser 2004). For an image with Gaussian noise convolved with a
Gaussian PSF with width $\sigma_g$ the density of peaks above a given
detection threshold, $\nu$ is given by 
\begin{equation}
n(>\nu) = \frac{1}{2^{5/2}\pi^{3/2}\sigma_g^{2}} \nu e^{-\nu^2 /2} 
\label{eq-theory}
\end{equation}
 To estimate the number of random
positive detections for a given image we must normalize this density
by the total number of realizations of the PSF within the image, i.e.,
\begin{equation}
N_{total}(>\nu) = n(\nu)*nrow*ncol/PSF_{FWHM}^2
\end{equation}
where nrow and ncol are the number of rows and columns in an images
and $PSF_{FWHM}$ is the FWHM of the PSF.

Figure~\ref{fig:peaks} shows good agreement in a comparison of this
model to a Monte-Carlo simulation for thresholds $4<\nu<6$ for a
4000x4072 pixel image with 0.88 arcsec seeing. The variance on these
measures is shown in Figure XXX {\bf Simon: I can make this figure on Monday morn, if needed.}
\begin{figure}
\includegraphics[width=0.5\textwidth,
height=0.25\textwidth]{figures/peaks.eps} 
\caption{A comparison of the predicted number of random peaks (solid
  line) detected in an image, as a function of significance, with
  those derived from a Monte-Carlo realization of a Gaussian noise
  image (solid points).}
\label{fig:peaks}
\end{figure}

@@

\subsection{Detection Threshold \label{sec-thresh}}
Obviously the number of false positives is strongly dependent on the detection threshold.  In order to
check the scaling of our false positive rates with detection threshold, we ran the same config (see Table~\ref{tab-bestconfig10} 
for the config parameters)
at several detection threshold values.  As discussed in 
Section~\ref{sec-analyticfp} one can predict the number of false positives expected in a sensor given
seeing and threshold values.  Values of the number of false positives (both +ve and -ve)
expected in LSST size detectors are given in Table~\ref{tab-fp}.

@@

In all cases the empirical results are below the predicted results indicating
that the number counts are smaller than the theoretical floor.  This is not entirely unexpected, as each
curve is strongly correlated, being a cumulative distribution, and represents only 9 samplings of the
background random Gaussian field.  We find that we can perform a 1--parameter correction to Equation \ref{eq-theory}
curves that
yields both the same slope and zeropoints as the theoretical curves, with the exception of the deconvolution 
visit.

Specifically, we define:
\begin{equation}
n\prime(>\nu) = n(>k\dot\nu)
\label{correction}
\end{equation}  
For our data we scale {\it down} $\nu$ in the model by $k = 0.975, 0.985 $~and~$ 0.99$ for 
v6866601, v6866602, and v6866603 respectively.  The data plotted with these corrections are shown
in the bottom row of Figure\ref{fig-fp_v_thresh}.  These results suggest our variance estimates 
are too {\it high}, and we are therefore detecting too {\it few} false detections.  This does not agree 
entirely with the analysis of the empirical variance, Section~\ref{sec-varmis} and Table~\ref{tab-variance1}, 
which suggest that our variance planes {\it underestimate} the true variance.  However, this makes clear that 
while we are at the level of the theoretical floor in false detection rate, it is imperative to track the 
variance in our data to at the 1\% level or better to optimize both the false detection rate, and the true detection
efficiency.

@@

Using this simplified data set, we can begin to look at how we may predict 
optimal configs based on measured statistics.  In particular, we investigate
the MSE, and show how it varies as a function of config in Figures~\ref{fig-MSE_FITTED} and \ref{fig-MSE_CONTROL}.
We find that the MSE of the fitted sample (Figure~\ref{fig-MSE_FITTED}) continually decreases as the complexity 
of the model increases.  However, for the control sample (Figure~\ref{fig-MSE_CONTROL}), a minimum is achieved in all cases.
This minimum value suggests the point where the model starts overfitting, and is itself a method
to suggest the optimal configuration for image differencing, along with false positive rate.
However, MSE of a control sample is a metric that may be applied to data where the truth is not known, and thus has
potential to be used in the nightly processing.
For more discussion on this see Section $8.11$ in
``The Second Best Book in the World'' (i.e. {Ivezi{\'c} et al., in press}).

@@

\begin{table}
\centering
\begin{tabular}{clcccc}
\hline
\multicolumn{6}{|c|}{Per Sensor MSE Comparison} \\
\hline
Visit    & Prefilter & N$_{Same}$ & N$_{<3}$ & $\Delta FP$ & max($\Delta FP$)\\
\hline
v6866601 & True      & 0      & 2        & 22 & 4\\
         & False     & 4      & 6        & 24 & 10\\
v6866602 & True      & 3      & 5        & 7 & 2\\
         & False     & 6      & 7        & 8 & 6\\
v6866603 & True      & 6      & 7        & 3 & 1\\
         & False     & 6      & 8        & 3 & 1\\
\end{tabular}
\caption{
We examine the configs chosen based on the minimum $\langle $MSE$\rangle$ and compare them to the config of the
minimum number of false positives.  The sample here consists of 9 sensors per visit.  
N$_{Same}$ is the number of times the minimum $\langle $MSE$\rangle$ coincides with the minimum in false positives.  N$_{<3}$
is the number of times the minimum false positive config is one of the top three configs when rank ordered by increasing
$\langle MSE \rangle$ value.  $\Delta FP$
is the total change in false positive count using the minimum $\langle $MSE$\rangle$ config as compared to the sum of the minimum 
false positives per sensor.  Column max($\Delta FP$) is the maximum change in detected false positives on any sensor in that visit.
\label{tab-MSE_persensor}}
\end{table}

@@

\begin{figure}%
\captionsetup[subfloat]{labelformat=empty}
\centering
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_01_nocorr.eps}}
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_02_nocorr.eps}}
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_03_nocorr.eps}}\\
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_01_corr.eps}}%
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_02_corr.eps}}%
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_03_corr.eps}}%
\caption{Here we show the comparison of detected false positive count to the analytic prediction.
In all cases the blue circles are the prefilter case and the red squares are the postfilter case.  
The top row has had the counts corrected for the underestimate of the noise by the variance plane (Table~\ref{tab-variance1}).  
In the bottom row we undertaken a 1--parameter correction to align the predicted and measured values.  See Section~\ref{sec-thresh}
for how this correction is applied.  The 2.5\%, 1.5\% and 1\% scaling in $k$ correspond to a 5\%, 3\%, 2\% over-estimation of the image variance, respectively
from Left to Right in the bottom pane.  
The shaded area is the 1-$\sigma$ confidence envelope
determined from Monte Carlo simulations assuming the analytic form of the false detection rate and a Gaussian distribution.}
\label{fig-fp_v_thresh}
\end{figure}
